Topological approach to finding blackholes in directed networks

### 2019-12-07 blackholes/pavt/article.tex
Small updates

\subsection{Known issues}
In this section we shortly discuss applicability of the iBlackhole algorithm, proposed in \cite{li2010detecting}.
The iBlackhole algorithm (see Algorithm \ref{alg:iblackhole}) is designed for searching blackholes of fixed size. Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which are quite
a small size.
Also, there are graphs, which has a small number of blackholes in relation to number of nodes. They consist
of several large SCCs and relatively small number of independent leaves (or roots). Here arise some questions considering applicability of
iBlackhole algorithm:
\begin{itemize}
\item How do we know if there are blackholes of certain size? We have to check all the possible sizes, which can take too much time even in parallel.
\item If we check blackhole sizes consecutively, what is the expected time of the first blackhole found? \FIXME{мы отвечаем на этот вопрос?}
\item How do we reduce the search space for the brute-force stage and avoid repeatedly checking the same combinations for different candidate sizes?
\end{itemize}

@@

\section{Algorithm Design}
We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
\FIXME{У нас есть ограничение на суммарное время фазу препроцессинга?}
The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
Decreased graph scale would definitely speed up the the calculation, as it can fall into a brute-force search in most cases.
The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in a restricted time.

@@

\begin{definition}
Topological sort or topological ordering of a directed graph is a linear ordering of 
its vertices such that for every directed edge $(u,v)$ from node $u$ to node $v$,
$v$ comes before $u$ in the ordering.
\end{definition}

@@

@@

Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
in preprocessing step, it still can result in lots of computation. It's worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
be filtered later in lazy manner.

Let's consider a closure of a single blackhole root. Let this closure to be acquired by depth-first search, which basically means, that
we can iterate over these nodes in topologically sorted order. Also, while traversing this graph, we can calculate size of closure for each node
in the given roots closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder[i]$. We will call such nodes special.

@@

\begin{table}[]
\caption{Experimental results}
\label{tabular:tableresults}
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c}
Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
\hline
RMAT 04 & chinese & 1 & 0.00039521 & topsort & 1 & 0.000115912 \\
RMAT & 06 & chinese & 1 & 0.0040142 & topsort & 1 & 0.00034812 \\
RMAT & 08 & chinese & 1 & 0.0637587 & topsort & 1 & 0.00146486 \\
RMAT & 10 & chinese & 6 & 1200 & topsort & 10 & 0.00684045 \\
RMAT & 12 & chinese & 29 & 1200 & topsort & 971738 & 1200 \\
RMAT & 14 & chinese & 187 & 1200 & topsort & 735128 & 1200 \\
RMAT & 16 & chinese & 1192 & 1200 & topsort & 373705 & 1200 \\
RMAT & 18 & chinese & 5161 & 1200 & topsort & 10286 & 1200 \\
RMAT & 20 & chinese & 609 & 1200 & topsort & 48867 & 1200 \\
RMAT & 22 & chinese & 108 & 1200 & topsort & 0 & 1200 \\
SSCA2 & 04 & chinese & 16 & 0.255101 & topsort & 16 & 0.000189925 \\
SSCA2 & 06 & chinese & 11 & 1200 & topsort & 70 & 0.00121528 \\
SSCA2 & 08 & chinese & 42 & 1200 & topsort & 263314 & 1200 \\
SSCA2 & 10 & chinese & 85 & 1200 & topsort & 114411 & 1200 \\
SSCA2 & 12 & chinese & 343 & 1200 & topsort & 5870 & 1200 \\
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Experimental results. Part 2}
\label{tabular:tableresults2}
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c}
Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
SSCA2 & 14 & chinese & 1433 & 1200 & topsort & 323671 & 1200 \\
SSCA2 & 16 & chinese & 5829 & 1200 & topsort & 893376 & 1200 \\
SSCA2 & 18 & chinese & 23179 & 1200 & topsort & 224567 & 1200 \\
SSCA2 & 20 & chinese & 92347 & 1200 & topsort & 24199 & 1200 \\
SSCA2 & 22 & chinese & 370770 & 1200 & topsort & 0 & 1200 \\
UR & 04 & chinese & 1 & 0.000263955 & topsort & 1 & 0.000115029 \\
UR & 06 & chinese & 1 & 0.00222731 & topsort & 1 & 0.000331235 \\
UR & 08 & chinese & 1 & 0.0868632 & topsort & 1 & 0.00147483 \\
UR & 10 & chinese & 1 & 1.93527 & topsort & 1 & 0.00670724 \\
UR & 12 & chinese & 1 & 46.1463 & topsort & 1 & 0.034011 \\
UR & 14 & chinese & 1 & 1169.28 & topsort & 1 & 0.47257 \\
UR & 16 & chinese & 0 & 1200 & topsort & 1 & 3.05003 \\
UR & 18 & chinese & 0 & 1200 & topsort & 1 & 14.5788 \\
UR & 20 & chinese & 0 & 1200 & topsort & 1 & 67.7176 \\
UR & 22 & chinese & 0 & 1200 & topsort & 1 & 373.244 \\
\end{tabular}
\end{center}
\end{table}
\pagebreak

@@

In this paper we formulated and addressed some issues in application of already known iBlackhole algorithm. We introduced and tested two different ideas of optimization, combined into
our TopSort algorithm. Those are: graph condensation preprocessing and topsort bruteforce optimization. We tested out approach on three different types of graphs and number of scales.
In case of large SCCs we demostrate definite advantage over iBlackhole algorithm. Then, we can relatively fast detect total absence of blackholes, but sometimes quite costly preprocessing
deminishes any boost of productivity and out performance degrades.

@@

\begin{figure}[p]
    \includegraphics[width=\linewidth]{rmat.pdf}
    \caption{RMAT}
    \label{fig:rmat}
\end{figure}
\begin{figure}[p]
    \includegraphics[width=\linewidth]{ssca2.pdf}
    \caption{SSCA2}
    \label{fig:ssca2}
\end{figure}
\begin{figure}[p]
    \includegraphics[width=\linewidth]{ur.pdf}
    \caption{UR}
    \label{fig:ur}
\end{figure}

>>>>

\subsection{Known issues}
In this section we shortly discuss applicability of the iBlackhole algorithm, proposed in \cite{li2010detecting}.
The iBlackhole algorithm (see Algorithm \ref{alg:iblackhole}) is designed for searching blackholes of fixed size. Potentially it has a great resource of parallelism, however this work does not cover the case of large scale graphs, as they only run their algorithm on 1500 node networks, which are quite
a small size.
Also, there are graphs, which has a small number of blackholes in relation to number of nodes. They consist
of several large SCCs and relatively small number of independent leaves (or roots). Here arise some questions considering applicability of
iBlackhole algorithm:
\begin{itemize}
\item How do we know if there are blackholes of certain size? We have to check all the possible sizes, which can take too much time even in parallel.
\item How do we reduce the search space for the brute-force stage and avoid repeatedly checking the same combinations for different candidate sizes?
\end{itemize}

@@

\section{Algorithm Design}
We propose to look at the blackhole search problem from the following: it would be divided into two independent tasks. 
\FIXME{ограничение на суммарное время - да, 20 минут}
The first one is graph preprocessing. Here we aim to simplify graph structure as far as it is possible with respect to a restricted time span. 
Decreased graph scale would definitely speed up the the calculation, as it can fall into a brute-force search in most cases.
The second is the blackhole search itself. This step is done after preprocessing and the goal here is to find as many blackholes as possible in a restricted time.

@@

\begin{definition}
Topological sort or topological ordering of a directed graph is a linear ordering of 
its vertices such that for every directed edge $(u,v)$ from node $u$ to node $v$,
$v$ comes before $u$ in the topological ordering.
\end{definition}

@@

\FIXME{избежать путаницы между brute-force в разных местах статьи}

@@

\FIXME{не совсем понял логику в общем тексте}
Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
in preprocessing step, it still can result in lots of computation. It is worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
be filtered later in lazy manner.

Let's consider a closure of a single blackhole root $r$. Let this closure to be acquired by depth-first search, which basically means, that
we can iterate by $i$ over these nodes in topologically sorted order $topsortOrder(r)[i]$. Also, while traversing this graph, we can calculate size of closure for each node
in the given root closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder(r)[i]$. We will call such nodes special.

@@

\begin{table}[]
\caption{Experimental results}
\label{tabular:tableresults}
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c}
Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
\hline
RMAT & 04 & NewJersey & 1 & 0.000265422 & TopSort & 1 & 0.000115912 \\
RMAT & 06 & NewJersey & 1 & 0.0121347 & TopSort & 1 & 0.00034812 \\
RMAT & 08 & NewJersey & 1 & 1.66685 & TopSort & 1 & 0.00146486 \\
RMAT & 10 & NewJersey & 6 & 1200 & TopSort & 10 & 0.00684045 \\
RMAT & 12 & NewJersey & 29 & 1200 & TopSort & 971738 & 1200 \\
RMAT & 14 & NewJersey & 187 & 1200 & TopSort & 735128 & 1200 \\
RMAT & 16 & NewJersey & 1192 & 1200 & TopSort & 373705 & 1200 \\
RMAT & 18 & NewJersey & 5176 & 1200 & TopSort & 10286 & 1200 \\
RMAT & 20 & NewJersey & 617 & 1200 & TopSort & 48867 & 1200 \\
RMAT & 22 & NewJersey & 98 & 1200 & TopSort & 0 & 1200 \\
SSCA2 & 04 & NewJersey & 16 & 0.154763 & TopSort & 16 & 0.000189925 \\
SSCA2 & 06 & NewJersey & 11 & 1200 & TopSort & 70 & 0.00121528 \\
SSCA2 & 08 & NewJersey & 42 & 1200 & TopSort & 263314 & 1200 \\
SSCA2 & 10 & NewJersey & 85 & 1200 & TopSort & 114411 & 1200 \\
SSCA2 & 12 & NewJersey & 343 & 1200 & TopSort & 5870 & 1200 \\
\end{tabular}
\end{center}
\end{table}

\begin{table}[]
\caption{Experimental results. Part 2}
\label{tabular:tableresults2}
\begin{center}
\begin{tabular}{c|c|c|c|c|c|c|c|c}
Graph & Type & Algorithm & Blackholes & Time(s) & Algorithm & Blackholes & Time(s) \\
SSCA2 & 14 & NewJersey & 1433 & 1200 & TopSort & 323671 & 1200 \\
SSCA2 & 16 & NewJersey & 5829 & 1200 & TopSort & 893376 & 1200 \\
SSCA2 & 18 & NewJersey & 23179 & 1200 & TopSort & 224567 & 1200 \\
SSCA2 & 20 & NewJersey & 92347 & 1200 & TopSort & 24199 & 1200 \\
SSCA2 & 22 & NewJersey & 370770 & 1200 & TopSort & 0 & 1200 \\
UR & 04 & NewJersey & 1 & 0.000223768 & TopSort & 1 & 0.000115029 \\
UR & 06 & NewJersey & 1 & 0.0170713 & TopSort & 1 & 0.000331235 \\
UR & 08 & NewJersey & 1 & 2.30256 & TopSort & 1 & 0.00147483 \\
UR & 10 & NewJersey & 1 & 240.822 & TopSort & 1 & 0.00670724 \\
UR & 12 & NewJersey & 0 & 1200 & TopSort & 1 & 0.034011 \\
UR & 14 & NewJersey & 0 & 1200 & TopSort & 1 & 0.47257 \\
UR & 16 & NewJersey & 0 & 1200 & TopSort & 1 & 3.05003 \\
UR & 18 & NewJersey & 0 & 1200 & TopSort & 1 & 14.5788 \\
UR & 20 & NewJersey & 0 & 1200 & TopSort & 1 & 67.7176 \\
UR & 22 & NewJersey & 0 & 1200 & TopSort & 1 & 373.244
\end{tabular}
\end{center}
\end{table}
\pagebreak

@@

In this paper we formulated and addressed some issues in application of already known iBlackhole algorithm. We introduced and tested two different ideas of optimization, combined into
our TopSort algorithm. Those are: graph condensation preprocessing and topsort bruteforce optimization. We tested out approach on three different types of graphs and number of scales.
In case of large SCCs we demostrate definite advantage over iBlackhole algorithm. Then, we can relatively fast detect total absence of blackholes, but sometimes quite costly preprocessing
deminishes any boost of productivity and out performance degrades.
\FIXME{можно написать в TODO}
If we check blackhole sizes consecutively, what is the expected time of the first blackhole found?

@@

\begin{figure}[p]
    \includegraphics[width=\linewidth]{rmat.png}
    \caption{RMAT}
    \label{fig:rmat}
\end{figure}
\begin{figure}[p]
    \includegraphics[width=\linewidth]{ssca2.png}
    \caption{SSCA2}
    \label{fig:ssca2}
\end{figure}
\begin{figure}[p]
    \includegraphics[width=\linewidth]{ur.png}
    \caption{UR}
    \label{fig:ur}
\end{figure}
